### 2016-09-03 Chapters/ContinuumSolvationModels.tex
Work on MQM chapter
Save the day's work. Chapter 1 almost done

How are quantum/classical polarizable Hamiltonians implemented in this
framework?
As shown in Chapter \ref{ch:QM} for the \acrshort{CC} method, one can
formulate methods in quantum chemistry as the unconstrained minimization
of a Lagrangian.\autocite{Helgaker1982-xl, Arponen1983-qt,
Helgaker1988-to, Helgaker1989-wl, Koch1990-vr, Kvaal2013-jr}
Let us call $\mathcal{E}(\vect{\eta})$ the definition of the energy
for the given quantum chemical method.
The \emph{amplitudes} $\vect{\eta}$ are determined by a set of
stationarity conditions $\mat{\Omega}(\vect{\eta}) = 0$.
Direct minimization of $\mathcal{E}(\vect{\eta})$ is thus a
\emph{constrained} optimization.
Introducing a set of Lagrange \emph{multipliers} $\bar{\vect{\eta}}$
lets us recast the problem as an unconstrained minimization of the
Lagrangian:
\begin{equation}
  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) =
  \mathcal{E}(\vect{\eta}) +
  \scalprod[W]{\bar{\vect{\eta}}}{\mat{\Omega}(\vect{\eta})}
\end{equation}
where the governing equations are obtained by differentiation with
respect to both sets of parameters:
\begin{subequations}
  \begin{align}
    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}}
    &= \mat{\Omega}(\vect{\eta}) = 0 \\
    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
    = 0
  \end{align}
\end{subequations}
Due to the  variational property of the classical polarizable functional
one can devise an \emph{effective} quantum/classical polarizable
Lagrangian:\autocite{Lipparini2016-mo}
\begin{equation}
  \mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p) =
  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) +
  \frac{1}{2}{}^t\p\V\p + {}^t\p\s(\vect{\eta}, \bar{\vect{\eta}})
\end{equation}
and derive the governing equations in the usual manner:
\begin{subequations}
  \begin{align}
    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\bar{\vect{\eta}}}
    &= \mat{\Omega}(\vect{\eta}) + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}} = 0 \\
    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\vect{\eta}}
    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
    + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
    = 0 \\
    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\p} &=
    \V\p + \s(\vect{\eta}, \bar{\vect{\eta}}) = 0
  \end{align}
\end{subequations}
The quantum mechanical and the classical model are thus fully coupled,
due to the dependence of the source term $\s$ on the amplitudes and
multipliers.

>>>>

How are quantum/classical polarizable Hamiltonians implemented in this
framework?
As shown in Chapter \ref{ch:QM} for the \acrshort{CC} method, one can
formulate methods in quantum chemistry as the unconstrained minimization
of a Lagrangian.
Due to the  variational property of the classical polarizable functional
one can devise an \emph{effective} quantum/classical polarizable
Lagrangian:\autocite{Lipparini2016-mo}
\begin{equation}
  \mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p) =
  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) +
  \frac{1}{2}{}^t\p\V\p + {}^t\p\s(\vect{\eta}, \bar{\vect{\eta}})
\end{equation}
and derive the governing equations in the usual manner:
\begin{subequations}
  \begin{align}
    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\bar{\vect{\eta}}}
    &= \mat{\Omega}(\vect{\eta}) + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}} = 0 \\
    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\vect{\eta}}
    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
    + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
    = 0 \\
    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\p} &=
    \V\p + \s(\vect{\eta}, \bar{\vect{\eta}}) = 0
  \end{align}
\end{subequations}
The quantum mechanical and the classical model are thus fully coupled,
due to the dependence of the source term $\s$ on the amplitudes and
multipliers.

\todo[inline]{Provide an example with single-reference theories.}
As an example, let us consider the single-reference \acrshort{HF} and
\acrshort{KS}-\acrshort{DFT} methods.

### 2016-09-03 Chapters/ContributionsSummary.tex
Save the day's work. Chapter 1 almost done

\lstinputlisting[language=C,
                 caption={Relevant calls to \pcmsolver in a C host program.
                 Ellipsis in the functions arguments list stand for omitted parameters.
                 For the full, functional example visit
                 \url{http://pcmsolver.readthedocs.io/en/latest/users/C-example.html}},
                 label={lst:C-host}
                ]{snippets/C_host.c}

@@

\lstinputlisting[language=Fortran,
                 caption={Relevant calls to \pcmsolver in a Fortran host program.
                 Notice the use of the standard Fortan ISO C bindings.
                 Ellipsis in the functions arguments list stand for omitted parameters.
                 For the full, functional example visit
                 \url{http://pcmsolver.readthedocs.io/en/latest/users/fortran-example.html}},
                 label={lst:Fortran-host}
                 ]{snippets/Fortran_host.F90}

>>>>

\lstinputlisting[language=C,
                 caption={Relevant calls to \pcmsolver in a C host program.
                 Ellipsis in the functions arguments list stand for omitted parameters.
                 For the full, functional example visit
                 \url{http://pcmsolver.readthedocs.io/en/v1.1.5/users/C-example.html}},
                 label={lst:C-host}
                ]{snippets/C_host.c}

@@

\lstinputlisting[language=Fortran,
                 caption={Relevant calls to \pcmsolver in a Fortran host program.
                 Notice the use of the standard Fortan ISO C bindings.
                 Ellipsis in the functions arguments list stand for omitted parameters.
                 For the full, functional example visit
                 \url{http://pcmsolver.readthedocs.io/en/v1.1.5/users/fortran-example.html}},
                 label={lst:Fortran-host}
                 ]{snippets/Fortran_host.F90}
